# Simulation Exercises on Exponential Distribution
========================================================

## Simulation

The first step is setting up the parameters for the simulation.
```{r vardec}
lambda <- 0.2
nsamples <- 40
nsims <- 1000
```

Now we run the simulations and store the resulting data.
A histogram is added as a quick summary of the resulting data
```{r runsims}
idx <- 0
means_data <- numeric()
set.seed(65455)
while(idx < nsims){
sim_mean <- mean(rexp(nsamples,lambda))
means_data <- c(means_data,sim_mean)
idx <- idx+1
}
```

## Review of simulation results

We can now review the resulting data.
```{r calc}
theoretical_mean <- 1/0.2
theoretical_variance <- (1/0.2)^2

sample_mean <- mean(means_data)
sample_variance <- var(means_data)
sample_sd <- sd(means_data)

expected_variance <- theoretical_variance/(nsamples)
```

We will use the ration of sampled value vs theoretical value to compare the results of the simulation.
A value of 1 indicates both values are equal. The further the ratio is from one, the bigger the difference between theoretical value and obtained value.

+ For the mean, the ratio of theoretical mean (`r theoretical_mean`) vs simulation mean (`r sample_mean`) is : `r sample_mean/theoretical_mean`.


+ For the variance, the ratio of expected theoretical variance (`r expected_variance`) vs simulation variance (`r sample_variance`) is : `r sample_variance/expected_variance`.

## Evaluation of distribution normality

```{r normalplot}
hist(means_data,prob=TRUE,col="orangered")
first_sd_volume <- length(means_data[means_data < sample_mean+sample_sd 
                                     & means_data > sample_mean-sample_sd])/nsims
second_sd_volume <- length(means_data[means_data < sample_mean+2*sample_sd 
                                      & means_data > sample_mean-2*sample_sd])/nsims
third_sd_volume <- length(means_data[means_data < sample_mean+3*sample_sd 
                                     & means_data > sample_mean-3*sample_sd])/nsims
```

It is expected that in a normal distribution, the percentage of observations within one, two and three standard deviation from the mean should be around 68%, 95% and 99% respectively.

The percentage of values within one, two, and three standard deviations from the mean in our simulation data is:
+ one SD : `r first_sd_volume`
+ two SD : `r second_sd_volume`
+ three SD : `r third_sd_volume`

## Evaluation of confidence interval

Finally we evaluate the confidence range for 1.96 SDs
```{r confidencerange}
cov_range <- 1.96 * sqrt(expected_variance) # sqrt(theoretical_variance/nsamples)
low_conf_value <- sample_mean - cov_range
high_conf_value <- sample_mean + cov_range
```

The confidence interval is : [`r low_conf_value`,`r high_conf_value`]

We can calculate the percentage of samples that are contained in this interval:
```{r rangecoverage}
coverage <- length(means_data[means_data > low_conf_value 
                              & means_data < high_conf_value])/nsims
```

The percentage of simulations that fall within the confidence range in these simulations is : `r coverage`